$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-04 paper/paper.tex
on contexts

    \item [Shared]
    The carrier is a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
    Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
    The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
    All other rules follow trivially.
    For example, a shared channel can be used unrestrictedly many times in either input or output.
\end{description}
\todo{Added a last for example sentence to all. check}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Typing Contexts}
\label{contexts}
\todo{unify typing system and type system. only use one terminology in the paper.}
\begin{nidefinition}
  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
  An indexed set of usage algebras is a set of indices $\Idx$ together with an interpretation $\Usage_{idx}$ of indices into types, and an interpretation of indices into usage algebras on that type.
  $\Idx$ must be non-empty.
  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:usage-algebra} for a formal definition.
\end{nidefinition}

We keep typing contexts ($\PreCtx$) and usage contexts ($\Ctx$) separate.
The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.

\begin{nidefinition}
  \autoref{fig:prectx} defines types as either unit types, base types or channel types.
  The unit type $\unit$ serves as a proof of inhabitance for types.
  The base type $\base{n}$ uses natural numbers as placeholders for types --- the host language can interpret the former into the latter.
  This is for convenience, so that there is less universe polymorphism polluting specifications.
  The type $\channel{t}{x}$ of a channel determines what type of data and what usage annotations are sent over that channel.
  We keep track of types in a length-indexed $\PreCtx_n$.

@@

    \inferrule
        {\Gamma : \Ctx_{idxs} \\ x : \Usage_{idx} ^{\func{2}}}
        {\Gamma \comma x : \Ctx_{idxs \comma idx}}
  \end{mathpar}
  \caption{Length-indexed context of carrier indices with a context of multiplicities on top.}
  \label{fig:ctx}
  \end{figure}
\end{nidefinition}

\begin{note}
  We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have a common underlying context of indices.
  All the properties of the partial monoid are trivially lifted.
\end{note}

>>>>

    \item [Shared]
    The carrier is a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
    Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
    The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
    All other rules follow trivially.
    For example, a shared channel can be used unrestrictedly many times in either input or output.
\end{description}
\todo{Added a last for example sentence to all. check}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Typing Contexts}
\label{contexts}
\todo{unify typing system and type system. only use one terminology in the paper.}
\begin{nidefinition}
  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
  An indexed set of usage algebras is a set of indices $\Idx$ together with an interpretation $\Usage_{idx}$ of indices into types, and an interpretation of indices into usage algebras on that type.
  $\Idx$ must be non-empty.
  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:usage-algebra} for a formal definition.
\end{nidefinition}
\todo{again, usage and idx occur here, but the formal defintion seems to be actually necessary. I'm thinking that both fig 12 and 13 are needed in the main text...}

We keep typing contexts ($\PreCtx$) and usage contexts ($\Ctx$) separate.
The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.
\todo{where is the def of CTX?}

\begin{nidefinition}
  \autoref{fig:prectx} defines types as either unit types (rule $\unit$), base types (rule $\base{n}$) or channel types (rule $\channel{t}{x}$).
  %
  The unit type $\unit$ serves as a proof of inhabitance for types.
  The base type $\base{n}$ uses natural numbers as placeholders for types --- the host language can interpret the former into the latter. \todo{I don't get this... give an example for int, and just talk about pi, rather than a host language.}
  %This is for convenience, so that there is less universe polymorphism polluting specifications.
  The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel. This notation is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output \cite{K07}.
  \todo{I don't really like the symbol between t and x. Why not a simple semicolon ; ?}
  We keep track of types in a length-indexed $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.

@@

    \inferrule
        {\Gamma : \Ctx_{idxs} \\ x : \Usage_{idx} ^{\func{2}}}
        {\Gamma \comma x : \Ctx_{idxs \comma idx}}
  \end{mathpar}
  \caption{Length-indexed context of carrier indices with a context of multiplicities on top.}
  \label{fig:ctx}
  \end{figure}
\end{nidefinition}
\todo{Maybe fig 6 can go in appendix in order to make room for 12 and 13 to come back? This whole paragraph from def 6 below could be cut and brought to app? and only commented upon}
\begin{note}
  We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have a common underlying context of indices.
  All the properties of the partial monoid are trivially lifted.
\end{note}
